---
title: "| Exploratory Data Analysis on \n| a Well Coordinates Dataset\n"
author: "Alfonso R. Reyes"
date: '`r format(Sys.time())`'
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: yes
  bookdown::html_document2:
    number_sections: no
    theme: readable
    toc: yes
  bookdown::tufte_html2:
    number_sections: no
    toc: yes
  bookdown::word_document2: null
fontsize: 12pt
linestretch: 1.0
link-citations: yes
---



```{r setup, include=FALSE, error=TRUE, message=FALSE, warning=FALSE} 

options(knitr.table.format = "html") 
knitr::opts_knit$set(eval.after = c('fig.cap', 'fig.scap'))
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r load_libraries, message=FALSE}
library(knitr)
library(kableExtra)
library(ggplot2)

library("poppr")
library("tidyverse")
```

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the Australia, France, Mexico, and the United States (Table \@ref(tab:isolate-table)). 


Table: (\#tab:MCG-aggressiveness) Mean aggressiveness ratings for the 10 most abundant MCG; groupings according to 95% family-wise confidence interval.

```{r MLG-ag, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
# MLHagg %>% agghux
knitr::kable(
  mtcars[1:6, 1:6])
```

Table: (\#tab:MCG-aggressiveness) Mean aggressiveness ratings for the 10 most abundant MCG; groupings according to 95% family-wise confidence interval.


Refer to Table \@ref(tab:MCG-aggressiveness).

```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
# read_csv("../../results/tables/mlg_table.csv", 
#          na = "NA", # spaces read in as spaces
#          col_types = cols(.default = col_character())) %>%
#   rename(`$P_{sex}$` = Psex) %>% 
#   rename(MLH = MLG) %>%
#   knitr::kable(format = "pandoc")

knitr::kable(head(iris, 15))
```

Table: (\#tab:mlg-table) The five most abundant Multilocus Haplotypes (MLH) with
the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.


The most abundant MLH was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five Regions (NE, MI, WA, CO, and ND). 


Nearly all populations showed evidence of linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 

```{r table_ia, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}

knitr::kable(head(mtcars))
```
Table: (\#tab:iatab) Genotypic diversity and Linkage Disequilibrium summary for geographic populations arranged by abundance after clone-correction by a hierarchy of
Region/Source/Host/Year. 


-----------------------------------------------------------------------------
Country   State  Field Code              Year Host                          N 
--------- ------ ---------- ----------------- -------------------------- ----
USA       CA     wmn               2004, 2005 Beryl, Bunsi, G122           18 

USA       CO     producer          2007, 2010 Pinto, Yellow                41 

                 wmn                     2003 GH                            1 

USA       ID     producer                2003 GH                            1 

USA       MI     wmn        2003, 2004, 2005, 11A, 37, 38, B07104,         43 
                                   2008, 2009 Beryl, Bunsi, Cornell,       
                                              G122, Orion, PO7863, WM31      

                 producer    2003, 2008, 2009 BL, Black, Fuji, GH,         19 
                                              Merlot, SR06233, unk,        
                                              Vista, Zorro 

USA       MN     wmn               2003, 2004 Beryl, Bunsi, G122           11 

USA       ND     producer          2007, 2010 unk                          53 

                 wmn                     2005 Beryl, Bunsi, G122            7 

USA       NE     wmn        2004, 2005, 2008, Beryl, Bunsi, G122,          27 
                                         2010 PO7683, unk                    

                 producer   2003, 2007, 2009, Beryl, Emerson, GH, Orion,   20 
                                         2010  Pinto, Weihing                 

USA       NY     producer                2003 GH                            1 

USA       OR     wmn               2003, 2004 Beryl, Bunsi, G122           15 

                 producer                2003 G122, GH                      2 

USA       WA     wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
                                         2008  Cornell, G122, Orion, PO7      
                                              104, PO7863, WM31               

                 producer          2003, 2007 GH, Merlot, Pinto, Redkid    23 

USA       WI     producer                2003 GH                            2 

Mexico    -      wmn                     2005 Beryl, Bunsi, G122           18 

France    -      wmn               2004, 2005 Beryl, Bunsi, G122           18 

                 producer                2012 unk                           4 

Australia -      wmn                     2004 Beryl, Bunsi, G122            4 

                 producer                2004 Beryl                         2 

-----------------------------------------------------------------------------

Table: (\#tab:isolate-table) Description of *Sclerotinia sclerotiorum* isolates used in this study.
N = Number of Isolates. 
Key abbreviations: wmn = white mold screening nursery, producer = producer field, unk = unknown cultivar.



