---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

https://hughjonesd.github.io/huxtable/huxtable.html

# Getting started
A huxtable is a way of representing a table of text data in R. You already know that R can represent a table of data in a data frame. For example, if  mydata is a data frame, then mydata[1, 2] represents the the data in row 1, column 2, and mydata$start_time is all the data in the column called start_time.

A huxtable is just a data frame with some extra properties. So, if myhux is a huxtable, then myhux[1, 2] represents the data in row 1 column 2, as before. But this cell will also have some other properties - for example, the font size of the text, or the colour of the cell border.

To create a table with huxtable, use the function huxtable, or hux for short. This works very much like data.frame.


## convert to huxtable

```{r}
library(huxtable)
library(dplyr)


```

```{r}
data(mtcars)
car_ht <- as_hux(mtcars)

# Select columns by name:
cars_mpg <- car_ht[, c("mpg", "cyl", "am")] 

# Order by number of cylinders:
cars_mpg <- cars_mpg[order(cars_mpg$cyl),]
head(cars_mpg)
```

## basic hux table

```{r}
cars_mpg <- cars_mpg                          %>% 
      huxtable::add_rownames(colname = "Car") %>% 
      huxtable::add_colnames()                %>%
      set_right_border(0.4)                   %>% 
      set_right_border_color("grey") %>% 
      set_caption('cars table')

# Show the first 5 rows:
cars_mpg[1:5,]
```

```{r}
car_ht <- car_ht                                          %>%
      huxtable::add_rownames(colname = "Car")             %>%
      slice(1:10)                                         %>% 
      select(Car, mpg, cyl, hp)                           %>% 
      arrange(hp)                                         %>% 
      filter(cyl > 4)                                     %>% 
      rename(MPG = mpg, Cylinders = cyl, Horsepower = hp) %>% 
      mutate(kml = MPG/2.82) 

car_ht
```


```{r}
car_ht <- car_ht                               %>% 
      set_number_format(1:7, "kml", 2)         %>% 
      set_col_width(c(.35, .15, .15, .15, .2)) %>% 
      set_width(.6)                            %>% 
      huxtable::add_colnames()                 %>% 
      set_right_border(0.4)                    %>% 
      set_right_border_color("grey")           %>% 
      set_caption('Formatted table')

car_ht
```

## 

```{r}
mtcars[1:4, 1:4]
```

```{r}
# a_hux removes rownames and colnames
as_hux(mtcars[1:4, 1:4])  
```


```{r}
as_hux(mtcars[1:4, 1:4])                           %>% 
      huxtable::add_rownames(colname = "Car name") %>% 
      huxtable::add_colnames() %>% 
    set_caption('Adding rows and columns')
```


```{r}
cars_mpg <- cbind(cylinders = cars_mpg$cyl, cars_mpg)
table(cars_mpg$cylinders) # how many rows for each group?
```

# Column and row spans
Sometimes you want a single cell to spread over more than one row or column: for example, if you want a heading that covers several different rows. You can do this using the colspan and rowspan properties. If you know HTML tables, then these will be familiar to you. colspan sets how many columns the cell covers, and rowspan sets how many rows the cell covers. Here, we’ll add some row and column headings to the mtcars dataset:

```{r}
cars_mpg$cylinders[1]   <- ""
cars_mpg$cylinders[2]   <- "Four cylinders"
cars_mpg$cylinders[13]  <- "Six cylinders"
cars_mpg$cylinders[20]  <- "Eight cylinders"

rowspan(cars_mpg)[2, 1]  <- 11
rowspan(cars_mpg)[13, 1] <- 7
rowspan(cars_mpg)[20, 1] <- 14

cars_mpg <- rbind(c("List of cars", "", "", "", ""), cars_mpg)
colspan(cars_mpg)[1, 1] <- 5
align(cars_mpg)[1, 1] <- "center"

# a little more formatting:

cars_mpg <- set_all_padding(cars_mpg, 2)
cars_mpg <- set_all_borders(cars_mpg, 1)
valign(cars_mpg)[1,] <- "top"
col_width(cars_mpg) <- c(.4 , .3 , .1, .1, .1)
number_format(cars_mpg)[, 4:5] <- 0
bold(cars_mpg)[1:2, ] <- TRUE
bold(cars_mpg)[, 1] <- TRUE
if (knitr::is_latex_output()) font_size(cars_mpg) <- 10
cars_mpg
```

```{r}
car_ht                                                 %>% 
      set_background_color(evens, everywhere, "wheat") %>% 
      set_background_color(odds, everywhere, grey(.9)) %>% 
      set_bold(1, everywhere, TRUE) %>% 
    set_caption('Set a background')
```


```{r}
car_ht %>% 
    set_background_color(everywhere, starts_with("C"), "orange")
```


```{r}
car_ht %>% set_italic(everywhere, matches("[aeiou]"), TRUE)
```


```{r}
head(mtcars)

```


```{r}
options(huxtable.knit_print_df = FALSE)

head(mtcars) # back to normal
```


# Quick output commands
Sometimes you quickly want to get your data into a Word, Excel, HTML or PDF document. To do this you can use the quick_docx, quick_xlsx,  quick_html and quick_pdf functions. These are called with one or more huxtable objects, or objects which can be turned into a huxtable such as data frames. A new document of the appropriate type will be created and opened. By default the file will be in the current directory under the name e.g. huxtable-output.pdf. If the file already exists, you’ll be asked to confirm the overwrite. For non-interactive use, you must specify a filename yourself explicitly – this keeps you from accidentally trashing your files.
```{r}
# quick_pdf(mtcars) 
quick_pdf(mtcars, file = 'motorcars data.pdf')
```

