---
title: "sclerotinia, isolate"
author: "Alfonso R. Reyes"
date: "Version of `r format(Sys.time())`"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: yes
  pdf_document:
    toc: yes
  bookdown::tufte_html2:
    number_sections: no
    toc: yes
  bookdown::word_document2: null
  bookdown::html_document2:
    number_sections: no
    theme: readable
    toc: yes
link-citations: yes
linestretch: 1
fontsize: 12pt
---




```{r setup, include=FALSE, error=TRUE, message=FALSE, warning=FALSE} 
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      error = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = 'center'
                      )
```

```{r}
library("tidyverse")
library("poppr")
```

```{r}
load("sclerotinia_16_loci.rda")

dat11cc  <- clonecorrect(dat11, ~Region/Source/Host/Year)
pop_size <- table(pop(dat11))

```


Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the Australia, France, Mexico, and the United States (Table \@ref(tab:isolate-table)). 
After collection, sclerotia were stored in Petri plates lined with filter paper, then stored at $20~^{\circ}$F or -$4~^{\circ}$C. 
Sclerotia were surface-sterilized with 50% Clorox bleach (at least 6% NaOCl, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O for 3 min.
The sterilized sclerotia were then placed on water agar plates (16g of Bacto agar per liter of ddH$_2$O, BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days. 
An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH$_2$O) [@otto-hanson2011variation].
In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].

# Results {-}

A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in 11 states in the United States as well as Australia, France, and Mexico (Table \@ref(tab:isolate-table)).
With the 11 loci used in the analyses (Table \@ref(tab:locus-stats)), we observed a total of 165 MLHs (215 with 16 loci).
These 11 loci are located on 7 chromosomes in the *S. sclerotiorum* genome with a minimum distance of 55Kbp between two loci on the same chromosome.
Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
Four regions had fewer than 10 isolates, Australia (`r pop_size[["Australia"]]`), WI (`r pop_size[["WI"]]`), NY (`r pop_size[["NY"]]`), ID (`r pop_size[["ID"]]`).
We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLHs (Fig. \@ref(biggraph)A,C).


# Supplementary Information {-}


<!-- This part is necessary for adding the "S" before supplementary figures.
Will not work with HTML outputs, though. -->

\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}


-----------------------------------------------------------------------------
Country   State  Field Code              Year Host                          N 
--------- ------ ---------- ----------------- -------------------------- ----
USA       CA     wmn               2004, 2005 Beryl, Bunsi, G122           18 

USA       CO     producer          2007, 2010 Pinto, Yellow                41 

                 wmn                     2003 GH                            1 

USA       ID     producer                2003 GH                            1 

USA       MI     wmn        2003, 2004, 2005, 11A, 37, 38, B07104,         43 
                                   2008, 2009 Beryl, Bunsi, Cornell,       
                                              G122, Orion, PO7863, WM31      

                 producer    2003, 2008, 2009 BL, Black, Fuji, GH,         19 
                                              Merlot, SR06233, unk,        
                                              Vista, Zorro 

USA       MN     wmn               2003, 2004 Beryl, Bunsi, G122           11 

USA       ND     producer          2007, 2010 unk                          53 

                 wmn                     2005 Beryl, Bunsi, G122            7 

USA       NE     wmn        2004, 2005, 2008, Beryl, Bunsi, G122,          27 
                                         2010 PO7683, unk                    

                 producer   2003, 2007, 2009, Beryl, Emerson, GH, Orion,   20 
                                         2010  Pinto, Weihing                 

USA       NY     producer                2003 GH                            1 

USA       OR     wmn               2003, 2004 Beryl, Bunsi, G122           15 

                 producer                2003 G122, GH                      2 

USA       WA     wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
                                         2008  Cornell, G122, Orion, PO7      
                                              104, PO7863, WM31               

                 producer          2003, 2007 GH, Merlot, Pinto, Redkid    23 

USA       WI     producer                2003 GH                            2 

Mexico    -      wmn                     2005 Beryl, Bunsi, G122           18 

France    -      wmn               2004, 2005 Beryl, Bunsi, G122           18 

                 producer                2012 unk                           4 

Australia -      wmn                     2004 Beryl, Bunsi, G122            4 

                 producer                2004 Beryl                         2 

-----------------------------------------------------------------------------

Table: (\#tab:isolate-table) Description of *Sclerotinia sclerotiorum* isolates used in this study.
N = Number of Isolates. 
Key abbreviations: wmn = white mold screening nursery, producer = producer field, unk = unknown cultivar.