---
title: "from sclerotinia"
author: "Alfonso R. Reyes"
date: '`r format(Sys.time())`'
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: yes
  bookdown::html_document2:
    number_sections: no
    theme: readable
    toc: yes
  bookdown::tufte_html2:
    number_sections: no
    toc: yes
  bookdown::word_document2: null
fontsize: 12pt
linestretch: 1.0
link-citations: yes
---




```{r load_libraries, message=FALSE}
library(knitr)
library(kableExtra)
library(ggplot2)

library("poppr")
library("tidyverse")
```



Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the Australia, France, Mexico, and the United States (Table \@ref(tab:isolate-table)). 

Refer to Table \@ref(tab:MCG-aggressiveness).

```{r MLG-ag, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
# MLHagg %>% agghux
knitr::kable(
  mtcars[1:6, 1:6])
```

Table: (\#tab:MCG-aggressiveness) Mean aggressiveness ratings for the 10 most abundant MCG; groupings according to 95% family-wise confidence interval.



```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
# read_csv("../../results/tables/mlg_table.csv", 
#          na = "NA", # spaces read in as spaces
#          col_types = cols(.default = col_character())) %>%
#   rename(`$P_{sex}$` = Psex) %>% 
#   rename(MLH = MLG) %>%
#   knitr::kable(format = "pandoc")

knitr::kable(head(iris, 15))
```

Table: (\#tab:mlg-table) The five most abundant Multilocus Haplotypes (MLH) with
the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.


The most abundant MLH was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five Regions (NE, MI, WA, CO, and ND). 


Nearly all populations showed evidence of linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 

```{r table_ia, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}

knitr::kable(head(mtcars))
```





