---
title: "huxtable, columns and rows span"
output:
  bookdown::pdf_document2:
    latex_engine: pdflatex
  html_notebook: default
---

```{r message=FALSE}
library(huxtable)
library(dplyr)
```


# Column and row spans
Sometimes you want a single cell to spread over more than one row or column: for example, if you want a heading that covers several different rows. You can do this using the colspan and rowspan properties. If you know HTML tables, then these will be familiar to you. colspan sets how many columns the cell covers, and rowspan sets how many rows the cell covers. Here, we’ll add some row and column headings to the mtcars dataset:

See Table \@ref(tab:colRowSpan)

See Table \@ref(colRowSpan)

```{r colRowSpan}
data(mtcars)
car_ht <- as_hux(mtcars)

# Select columns by name:
cars_mpg <- car_ht[, c("mpg", "cyl", "am")] 

# Order by number of cylinders:
cars_mpg <- cars_mpg[order(cars_mpg$cyl),]

cars_mpg <- cars_mpg                          %>% 
      huxtable::add_rownames(colname = "Car") %>% 
      huxtable::add_colnames()                %>%
      set_right_border(0.4)                   %>% 
      set_right_border_color("grey") %>% 
      set_caption('cars table')

cars_mpg <- cbind(cylinders = cars_mpg$cyl, cars_mpg)
# cars_mpg

cars_mpg$cylinders[1]   <- ""
cars_mpg$cylinders[2]   <- "Four cylinders"
cars_mpg$cylinders[13]  <- "Six cylinders"
cars_mpg$cylinders[20]  <- "Eight cylinders"

rowspan(cars_mpg)[2, 1]  <- 11
rowspan(cars_mpg)[13, 1] <- 7
rowspan(cars_mpg)[20, 1] <- 14

cars_mpg <- rbind(c("List of cars", "", "", "", ""), cars_mpg)
colspan(cars_mpg)[1, 1] <- 5
align(cars_mpg)[1, 1] <- "center"

# a little more formatting:

cars_mpg <- set_all_padding(cars_mpg, 2)
cars_mpg <- set_all_borders(cars_mpg, 1)
valign(cars_mpg)[1,] <- "top"
col_width(cars_mpg) <- c(.4 , .3 , .1, .1, .1)
number_format(cars_mpg)[, 4:5] <- 0
bold(cars_mpg)[1:2, ] <- TRUE
bold(cars_mpg)[, 1] <- TRUE
if (knitr::is_latex_output()) font_size(cars_mpg) <- 10
cars_mpg
```


# another way of printing. Second

See Table \@ref(tab:colRowSpan2)


```{r colRowSpan2, echo=FALSE}
data(mtcars)
car_ht <- as_hux(mtcars)

# Select columns by name:
cars_mpg <- car_ht[, c("mpg", "cyl", "am")] 

# Order by number of cylinders:
cars_mpg <- cars_mpg[order(cars_mpg$cyl),]

cars_mpg <- cars_mpg                          %>% 
      huxtable::add_rownames(colname = "Car") %>% 
      huxtable::add_colnames()                %>%
      set_right_border(0.4)                   %>% 
      set_right_border_color("grey") %>% 
      set_caption('cars table')

cars_mpg <- cbind(cylinders = cars_mpg$cyl, cars_mpg)
# cars_mpg

cars_mpg$cylinders[1]   <- ""
cars_mpg$cylinders[2]   <- "Four cylinders"
cars_mpg$cylinders[13]  <- "Six cylinders"
cars_mpg$cylinders[20]  <- "Eight cylinders"

rowspan(cars_mpg)[2, 1]  <- 11
rowspan(cars_mpg)[13, 1] <- 7
rowspan(cars_mpg)[20, 1] <- 14

cars_mpg <- rbind(c("List of cars", "", "", "", ""), cars_mpg)
colspan(cars_mpg)[1, 1] <- 5
align(cars_mpg)[1, 1] <- "center"

# a little more formatting:

cars_mpg <- set_all_padding(cars_mpg, 2)
cars_mpg <- set_all_borders(cars_mpg, 1)
valign(cars_mpg)[1,] <- "top"
col_width(cars_mpg) <- c(.4 , .3 , .1, .1, .1)
number_format(cars_mpg)[, 4:5] <- 0
bold(cars_mpg)[1:2, ] <- TRUE
bold(cars_mpg)[, 1] <- TRUE
if (knitr::is_latex_output()) font_size(cars_mpg) <- 10
cars_mpg
```

Table: (\#tab:colRowSpan2) Summary of _Categorical_ variables for the original dataset.


# another way of printing. Third

See Table \@ref(colRowSpan3)


```{r colRowSpan3, echo=FALSE}
data(mtcars)
car_ht <- as_hux(mtcars)

# Select columns by name:
cars_mpg <- car_ht[, c("mpg", "cyl", "am")] 

# Order by number of cylinders:
cars_mpg <- cars_mpg[order(cars_mpg$cyl),]

cars_mpg <- cars_mpg                          %>% 
      huxtable::add_rownames(colname = "Car") %>% 
      huxtable::add_colnames()                %>%
      set_right_border(0.4)                   %>% 
      set_right_border_color("grey") # %>% 
      # set_caption('cars table')

cars_mpg <- cbind(cylinders = cars_mpg$cyl, cars_mpg)
# cars_mpg

cars_mpg$cylinders[1]   <- ""
cars_mpg$cylinders[2]   <- "Four cylinders"
cars_mpg$cylinders[13]  <- "Six cylinders"
cars_mpg$cylinders[20]  <- "Eight cylinders"

rowspan(cars_mpg)[2, 1]  <- 11
rowspan(cars_mpg)[13, 1] <- 7
rowspan(cars_mpg)[20, 1] <- 14

cars_mpg <- rbind(c("List of cars", "", "", "", ""), cars_mpg)
colspan(cars_mpg)[1, 1] <- 5
align(cars_mpg)[1, 1] <- "center"

# a little more formatting:

cars_mpg <- set_all_padding(cars_mpg, 2)
cars_mpg <- set_all_borders(cars_mpg, 1)
valign(cars_mpg)[1,] <- "top"
col_width(cars_mpg) <- c(.4 , .3 , .1, .1, .1)
number_format(cars_mpg)[, 4:5] <- 0
bold(cars_mpg)[1:2, ] <- TRUE
bold(cars_mpg)[, 1] <- TRUE
if (knitr::is_latex_output()) font_size(cars_mpg) <- 10
cars_mpg
```

Table: (\#colRowSpan3) Summary of _Categorical_ variables for the original dataset.






(ref:colRowSpan4) Categorical variables summary.

```{r colRowSpan4, echo=FALSE}
data(mtcars)
car_ht <- as_hux(mtcars)

# Select columns by name:
cars_mpg <- car_ht[, c("mpg", "cyl", "am")] 

# Order by number of cylinders:
cars_mpg <- cars_mpg[order(cars_mpg$cyl),]

cars_mpg <- cars_mpg                          %>% 
      huxtable::add_rownames(colname = "Car") %>% 
      huxtable::add_colnames()                %>%
      set_right_border(0.4)                   %>% 
      set_right_border_color("grey") %>% 
      set_caption('(ref:colRowSpan4) ')

cars_mpg <- cbind(cylinders = cars_mpg$cyl, cars_mpg)
# cars_mpg

cars_mpg$cylinders[1]   <- ""
cars_mpg$cylinders[2]   <- "Four cylinders"
cars_mpg$cylinders[13]  <- "Six cylinders"
cars_mpg$cylinders[20]  <- "Eight cylinders"

rowspan(cars_mpg)[2, 1]  <- 11
rowspan(cars_mpg)[13, 1] <- 7
rowspan(cars_mpg)[20, 1] <- 14

cars_mpg <- rbind(c("List of cars", "", "", "", ""), cars_mpg)
colspan(cars_mpg)[1, 1] <- 5
align(cars_mpg)[1, 1] <- "center"

# a little more formatting:

cars_mpg <- set_all_padding(cars_mpg, 2)
cars_mpg <- set_all_borders(cars_mpg, 1)
valign(cars_mpg)[1,] <- "top"
col_width(cars_mpg) <- c(.4 , .3 , .1, .1, .1)
number_format(cars_mpg)[, 4:5] <- 0
bold(cars_mpg)[1:2, ] <- TRUE
bold(cars_mpg)[, 1] <- TRUE
if (knitr::is_latex_output()) font_size(cars_mpg) <- 10
cars_mpg

# Table: (\#tab:colRowSpan4) Summary of _Categorical_ variables for the original dataset.
```



# another way of printing. Fourth

See Table \@ref(tab:colRowSpan4)

See Table \@ref(colRowSpan4)

See Table \#ref(colRowSpan4)

See Table \#ref(tab:colRowSpan4)
