---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

https://hughjonesd.github.io/huxtable/huxtable.html

# Getting started
A huxtable is a way of representing a table of text data in R. You already know that R can represent a table of data in a data frame. For example, if  mydata is a data frame, then mydata[1, 2] represents the the data in row 1, column 2, and mydata$start_time is all the data in the column called start_time.

A huxtable is just a data frame with some extra properties. So, if myhux is a huxtable, then myhux[1, 2] represents the data in row 1 column 2, as before. But this cell will also have some other properties - for example, the font size of the text, or the colour of the cell border.

To create a table with huxtable, use the function huxtable, or hux for short. This works very much like data.frame.

```{r}
library(huxtable)


ht <- hux(
        Employee = c('John Smith', 'Jane Doe', 'David Hugh-Jones'),
        Salary = c(50000, 50000, 40000),
        add_colnames = TRUE
      )

ht
```



```{r}

bold(ht)[1,]           <- TRUE
bottom_border(ht)[1,]  <- 2
align(ht)[,2]          <- 'right'
right_padding(ht)      <- 10
left_padding(ht)       <- 10
width(ht)              <- 0.35
number_format(ht)      <- 2

ht
```

```{r}
tribble_hux(
  ~Employee,          ~Salary,
  "John Smith",       50000,
  "Jane Doe",         50000,
  "David Hugh-Jones", 40000,
  add_colnames = TRUE
)
```

```{r}
print_screen(ht)     # on the R command line, you can just type "ht"

```

```{r}
ht
```

```{r}
right_padding(ht) <- 10
left_padding(ht)  <- 10
```


```{r}
number_format(ht) <- 2    # 2 decimal places
```


```{r}
bold(ht)[1, ]          <- TRUE
bottom_border(ht)[1, ] <- 1
```

```{r}
ht
```


So far, all these properties have been set at cell level. Different cells can have different alignment, text formatting and so on. By contrast, caption is a table-level property. It only takes one value, which sets a table caption.

```{r}
caption(ht) <- 'Employee table'
ht
```



# Tidyverse syntax
If you prefer a tidyverse style of code, using the pipe operator %>%, then you can use set_* functions to set properties. These have the same name as the property, with set_ prepended. For example, to set the bold property, you use the set_bold function.

set_* functions return the modified huxtable, so you can chain them together like this:

```{r}
library(dplyr)
hux(
        Employee     = c('John Smith', 'Jane Doe', 'David Hugh-Jones'), 
        Salary       = c(50000, 50000, 40000),
        add_colnames = TRUE
      )                               %>%
      set_right_padding(10)           %>%
      set_left_padding(10)            %>% 
      set_bold(1, 1:2, TRUE)          %>% 
      set_bottom_border(1, 1:2, 1)    %>%
      set_align(1:4, 2, 'right')      %>%
      set_number_format(2)            %>% 
      set_caption('Employee table')
```


## Getting properties
```{r}
italic(ht)
```

```{r}
bottom_border(ht)[1:2,]
```


# Changing text in a huxtable

```{r}
ht[3, 1] <- 'Jane Jones'
ht
```


```{r}
ht_with_roles <- ht
ht_with_roles$Role <- c("Role", "Admin", "CEO", "Dogsbody")
ht_with_roles
```

```{r}
ht_with_roles <- cbind(ht, c("Role", "Admin", "CEO", "Dogsbody"))
ht_with_roles
```

```{r}
rbind(ht, c("Yihui Xie", 100000))

```


```{r}
rbind(ht, c("Yihui Xie", 100000), copy_cell_props = FALSE)

```

```{r}
ht <- insert_row(ht, "Hadley Wickham", "100000", after = 3)
ht <- add_footnote(ht, "DHJ deserves a pay rise")
ht
```


# Automatic formatting
By default, when you create a huxtable using huxtable or as_huxtable, the package will guess defaults for number formatting and alignment, based on the type of data in your columns. Numeric data will be right-aligned or aligned on the decimal point; character data will be left aligned; and the package will try to set sensible defaults for number formatting. If you want to, you can turn this off with autoformat = FALSE:


```{r}
my_data <- data.frame(
        Employee           = c("John Smith", "Jane Doe", "David Hugh-Jones"), 
        Salary             = c(50000L, 50000L, 40000L),
        Performance_rating = c(8.9, 9.2, 7.8)  
      )
as_huxtable(my_data, add_colnames = TRUE) # with automatic formatting
```

```{r}
as_huxtable(my_data, add_colnames = TRUE, autoformat = FALSE) # no automatic formatting

```


## Width and cell wrapping

```{r}
width(ht) <- 0.35
col_width(ht) <- c(.7, .3)
ht
```

```{r}
ht_wrapped <- ht
ht_wrapped[5, 1] <- "David Arthur Shrimpton Hugh-Jones"
wrap(ht_wrapped) <- TRUE
ht_wrapped
```

