---
title: "R Notebook"
output:
  pdf_document: 
    latex_engine: lualatex
  html_notebook: default
  html_document: default
---

```{r}

```


```{r bunch_o_figs, fig.height=4, fig.width=8}
n <- 100
x <- rnorm(n)
par(mfrow=c(1,2), las=1)
for(i in 1:8) {
  y <- i*x + rnorm(n)
  plot(x, y, main=i)
}
```
For R Markdown, the default graphics device is png. You can choose a different device using the chunk option dev. For example, to use the svg device (for Scalable Vector Graphics, which may look better in a web page as they can be scaled without loss of quality), you would use dev='svg', as follows:

```{r bunch_o_figs_svg, fig.height=4, fig.width=8, dev='svg'}
n <- 100
x <- rnorm(n)
par(mfrow=c(1,2), las=1)
for(i in 1:8) {
  y <- i*x + rnorm(n)
  plot(x, y, main=i)
}
```
You can pass arguments to the graphics device using the chunk option dev.args, which takes a list. For example:

```{r bunch_o_figs_pointsize, fig.height=4, fig.width=8, dev.args=list(pointsize=18)}
n <- 100
x <- rnorm(n)
par(mfrow=c(1,2), las=1)
for(i in 1:8) {
  y <- i*x + rnorm(n)
  plot(x, y, main=i)
}
```


The chunk option dev.args=list(pointsize=18) passes pointsize=18 to the png() device. This is equivalent to, in R, using png("myfile.png", pointsize=18).

# Tables

I try to avoid tables; figures are almost always better. And for informal reports, I’ll often just print out a matrix or data frame, rather than create a formal table.
## kable

If you want to make a somewhat nicer table, the simplest approach is to use the kable function in the knitr package. It doesn’t have many options, but in many cases it’s sufficient. Here’s an example.

```{r kable}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)
library(knitr)
kable(summary(out)$coef, digits=2)
```





