---
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

http://rapporter.github.io/pander/knitr.html

```{r setup, include=FALSE, error=TRUE, message=FALSE, warning=FALSE} 
knitr::opts_chunk$set(echo=TRUE, 
                      comment=NA, 
                      error=TRUE, 
                      warning=FALSE, 
                      message = FALSE, 
                      fig.align = 'center')
```

```{r}
knitr::kable(head(iris))
```

```{r}
library(pander)
library(descr, quietly = TRUE)
ct <- CrossTable(mtcars$gear, mtcars$cyl)

knitr::kable(ct)


```

```{r}
library(tables, quietly = TRUE)
```

```{r}
tab <- tabular( (Species + 1) ~ (n=1) + Format(digits=2)*
         (Sepal.Length + Sepal.Width)*(mean + sd), data=iris )
knitr::kable(tab)
```

```{r}
library(descr, quietly = TRUE)
pander(CrossTable(mtcars$gear, mtcars$cyl))
```


```{r}
library(tables, quietly = TRUE)
tab <- tabular( (Species + 1) ~ (n=1) + Format(digits=2)*
         (Sepal.Length + Sepal.Width)*(mean + sd), data=iris )
pander(tab)
```

If you don’t want the results of pander to be converted automatically, just set knitr.auto.asis to FALSE using panderOptions:

```{r}
panderOptions('knitr.auto.asis', FALSE)
pander(head(iris))
```

```{r}
panderOptions('knitr.auto.asis', TRUE)
```


As you can see, this doesn’t work correctly because pander tries to return a knit_asis class object when run inside knitr, but for loops/vectorized functions this results in incorrect output. The recommended way to solve this is to disable this behavior by setting knitr.auto.asis to FALSE using panderOptions. However, we also need to tell knitr to convert the table on the fly by specifying results='asis' in the chunk options:

```{r}
panderOptions('knitr.auto.asis', FALSE)
dfs <- list(mtcars[1:3, 1:4], mtcars[4:6, 1:4], mtcars[7:9, 1:4])
invisible(lapply(dfs, pander))
```

