---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE, error=TRUE, message=FALSE, warning=FALSE} 
knitr::opts_chunk$set(echo=TRUE, 
                      comment=NA, 
                      error=TRUE, 
                      warning=FALSE, 
                      message = FALSE, 
                      results = "asis",
                      fig.align = 'center')
```

```{r}
library(pander)
```


```{r}
pandoc.table(mtcars[1:3, 1:4])
```


```{r}
m <- mtcars[1:2, 1:3]
pandoc.table(m)

```


```{r}
pandoc.table(m, style = "grid")

```


```{r}
pandoc.table(m, style = "rmarkdown")
```

# Caption
It's really easy to add a caption to a table:

```{r}
pandoc.table(m, style = "grid", caption = "Hello caption!")
```


```{r}
set.caption("Hello caption!")
pandoc.table(m)
```

Unless permanent option is set for TRUE (by default), caption will be set only for next table. To disable permanently set caption, just call set.caption(NULL) or call set.caption with permanent parameter being set to FALSE.

# Highlighting cells
One of the fanciest features in pander is the ease of highlighting rows, columns or any cells in a table. This is a real markdown feature without custom HTML or LaTeX-only tweaks, so all HTML/PDF/MS Word/OpenOffice etc. formats are supported.

This can be achieved by calling pandoc.table directly and passing any (or more) of the following arguments or calling the R function with the same names before rendering a table with either the pander generic S3 method or via pandoc.table:

```{r}
t <- mtcars[1:3, 1:5]
emphasize.italics.cols(1)
emphasize.italics.rows(1)
emphasize.strong.cells(which(t > 20, arr.ind = TRUE))
pandoc.table(t)
```


```{r}
pandoc.table(head(iris[,1:3], 2), justify = c('right', 'center', 'left'))
```

```{r}
set.alignment('left', row.names = 'right')

```

