---
title: "pander example"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, error=TRUE, message=FALSE, warning=FALSE} 
knitr::opts_chunk$set(echo=TRUE, 
                      comment=NA, 
                      error=TRUE, 
                      warning=FALSE, 
                      message = FALSE, 
                      fig.align = 'center')
```

```{r}
library(pander)

data(iris)
knitr::kable(head(iris))
```


```{r}
irisSummary <- apply(iris[, 1:4], 2, function(x) tapply(x, iris$Species,
summary))
irisSummary <- lapply(irisSummary, do.call, what = rbind)

irisSummary
```


Each of these can again be displayed as a nicely formatted table using kable but unfortunately information
about the column that was summarised will be lost in the process. For some output formats kable supports
the use of a caption option but unfortuantely this doesn’t work when producing Markdown, as is the case
here. An alternative is to use the pander package to produce output suitable for further processing with
pandoc.


```{r}
suppressPackageStartupMessages(library(pander))
panderOptions("table.split.table", Inf) ## don't split tables
pander(irisSummary[1:3])
```

```{r}
for (i in 3:4) {
set.caption(sub(".", " ", names(irisSummary)[i], fixed = TRUE))
pander(irisSummary[[i]])
}
```

```{r}
pander(t.test(Sepal.Length ~ Species == "setosa", data = iris))
```

