---
title: "R Notebook"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: yes
  html_notebook: default
---

```{r setup, include=FALSE, error=TRUE, message=FALSE, warning=FALSE} 
knitr::opts_chunk$set(echo=TRUE, 
                      comment=NA, 
                      error=TRUE, 
                      warning=FALSE, 
                      message = FALSE, 
                      results = "asis",
                      fig.align = 'center')
```

https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html

```{r}
set.seed(42)
library(dplyr)
## 
## Attaching package: 'dplyr'
## The following objects are masked from 'package:stats':
## 
##     filter, lag
## The following objects are masked from 'package:base':
## 
##     intersect, setdiff, setequal, union
library(qwraps2)

# define the markup language we are working in.
# options(qwraps2_markup = "latex") is also supported.
options(qwraps2_markup = "markdown")

data(mtcars)

mtcars2 <-
  dplyr::mutate(mtcars,
                cyl_factor = factor(cyl,
                                    levels = c(6, 4, 8),
                                    labels = paste(c(6, 4, 8), "cylinders")),
                cyl_character = paste(cyl, "cylinders"))

str(mtcars2)
```

```{r}
with(mtcars2, table(cyl_factor, cyl_character))
##              cyl_character
## cyl_factor    4 cylinders 6 cylinders 8 cylinders
##   6 cylinders           0           7           0
##   4 cylinders          11           0           0
##   8 cylinders           0           0          14
with(mtcars2, all.equal(factor(cyl_character), cyl_factor))
## [1] "Attributes: < Component \"levels\": 2 string mismatches >"
```

```{r}
mean_sd(mtcars2$mpg)
## [1] "20.09 &plusmn; 6.03"
mean_sd(mtcars2$mpg, denote_sd = "paren")
## [1] "20.09 (6.03)"
```

```{r}
mci <- mean_ci(mtcars2$mpg)
mci
## [1] "20.09 (18.00, 22.18)"
print(mci, show_level = TRUE)
## [1] "20.09 (95% CI: 18.00, 22.18)"
```

```{r}
median_iqr(mtcars2$mpg)
## [1] "19.20 (15.43, 22.80)"
```

```{r}
n_perc(mtcars2$cyl == 4)
## [1] "11 (34.38%)"
n_perc0(mtcars2$cyl == 4)
## [1] "11 (34)"

n_perc(mtcars2$cyl_factor == 4)  # this returns 0 (0.00%)
## [1] "0 (0.00%)"
n_perc(mtcars2$cyl_factor == "4 cylinders")
## [1] "11 (34.38%)"
n_perc(mtcars2$cyl_factor == levels(mtcars2$cyl_factor)[2])
## [1] "11 (34.38%)"

# The count and percentage of 4 or 6 cylinders vehicles in the data set is
n_perc(mtcars2$cyl %in% c(4, 6))
## [1] "18 (56.25%)"
```

```{r}
x <- runif(6, min = 4, max = 70)

# geometric mean
mu_g <- prod(x) ** (1 / length(x))
mu_g
## [1] 46.50714
exp(mean(log(x)))
## [1] 46.50714
1.2 ** mean(log(x, base = 1.2))
## [1] 46.50714

# geometric standard deviation
exp(sd(log(x)))  ## This is wrong
## [1] 1.500247

# these equations are correct
sigma_g <- exp(sqrt(sum(log(x / mu_g) ** 2) / length(x)))
sigma_g
## [1] 1.448151

exp(sqrt((length(x) - 1) / length(x)) * sd(log(x)))
## [1] 1.448151
```

# Building a Data Summary Table
Objective: build a table reporting summary statistics for some of the variables in the mtcars2 data.frame overall and within subgroups. We’ll start with something very simple and build up to something bigger.

Let’s report the min, max, and mean (sd) for continuous variables and n (%) for categorical variables. We will report mpg, disp, wt, and gear overall and by number of cylinders.

The function summary_table, along with some dplyr functions will do the work for us. summary_table takes two arguments:

.data a (grouped_df) data.frame
summaries a list of summaries. This is a list-of-lists. The outer list defines the row groups and the inner lists define the specif summaries.

## overall, mtcars2, oursummary1

See Table \@ref(tab:ourSummary).

```{r }
our_summary1 <-
  list("Miles Per Gallon" =
       list("min" = ~ min(mpg),
            "max" = ~ max(mpg),
            "mean (sd)" = ~ qwraps2::mean_sd(mpg)),
       "Displacement" =
       list("min" = ~ min(disp),
            "max" = ~ max(disp),
            "mean (sd)" = ~ qwraps2::mean_sd(disp)),
       "Weight (1000 lbs)" =
       list("min" = ~ min(wt),
            "max" = ~ max(wt),
            "mean (sd)" = ~ qwraps2::mean_sd(wt)),
       "Forward Gears" =
       list("Three" = ~ qwraps2::n_perc0(gear == 3),
            "Four"  = ~ qwraps2::n_perc0(gear == 4),
            "Five"  = ~ qwraps2::n_perc0(gear == 5))
       )

```


```{r ourSummary, results='asis'}
### Overall
summary_table(mtcars2, our_summary1)
```

Table: (\#tab:ourSummary) mtcars2. Our Summary #1


## By number of cylinders
See Table \@ref(tab:cylFactor).

```{r cylFactor}
### By number of Cylinders
summary_table(dplyr::group_by(mtcars2, cyl_factor), our_summary1)
```

Table: (\#tab:cylFactor) mtcars2. By number of cylinders


## Change column names
If you want to change the column names, do so via the cnames argument to qable via the print method for `qwraps2_summary_table` objects. Any argument that you want to send to qable can be sent there when explicitly using the print method for `qwraps2_summary_table` objects.
See Table \@ref(tab:changeColumnNames).

```{r}
print(summary_table(dplyr::group_by(mtcars2, cyl_factor), our_summary1),
      rtitle = "Summary Statistics",
      cnames = c("Col 1", "Col 2", "Col 3"))
```

Table: (\#tab:changeColumnNames) If you want to change the column names, do so via the `cnames` argument

# Easy building of the summaries
The task of building the summaries list-of-lists can be tedious. tab_summary is designed to make it easier. For numeric variables, tab_summary will provide the formulae for the min, median (iqr), mean (sd), and max. factor and  character vectors will have calls to qwraps2::n_perc for all levels provided.

For version 0.2.3.9000 or beyond, arguments have been added to tab_summary to help control some of the formatting of counts and percentages. The original behavior of tab_summary used n_perc0 to format the summary of categorical variables. Now, n_perc is called and the end user can specify formatting options via a list passed via the n_perc_args argument. The default settings for tab_summary is below.


```{r}
args(tab_summary)
```

```{r}
tab_summary(mtcars2$mpg)

```


```{r}
tab_summary(mtcars2$gear) # gear is a numeric vector!

```


```{r}
tab_summary(factor(mtcars2$gear))

```

The our_summary1 object can be recreated as follows. Some additional row groups are provided to show default behavior of tab_summary. Important: Note that the tab_summary are made while using with. Further explanation for this follows.
See Table \@ref(tab:usingTabSummary).

```{r}
our_summary2 <-
  with(mtcars2,
       list("Miles Per Gallon" = tab_summary(mpg)[c(1, 4, 3)],
            "Displacement (default summary)" = tab_summary(disp),
            "Displacement" = c(tab_summary(disp)[c(1, 4, 3)],
                               "mean (95% CI)" = ~ frmtci(qwraps2::mean_ci(disp))),
            "Weight (1000 lbs)" = tab_summary(wt)[c(1, 4, 3)],
            "Forward Gears" = tab_summary(as.character(gear))
            ))
```


```{r}
whole <- summary_table(mtcars2, our_summary2)
whole
```

Table: (\#tab:usingTabSummary) A **whole** table. Note that the tab_summary are made while using `with()`


## Group by multiple factors:
See Table \@ref(tab:multipleFactors).

```{r}
grouped <- summary_table(dplyr::group_by(mtcars2, am, vs), our_summary2)
grouped
```

Table: (\#tab:multipleFactors) A **grouped** table. Use `dplyr::group_by` to group by multiple factors

## As one table:
See Table \@ref(tab:oneTable) 

```{r}
both <- cbind(whole, grouped)
both
```

Table: (\#tab:oneTable) A **one** table. Using `cbind` to bind two tables `whole` and `grouped`


## Adding P-values to a Summary Table
There are many, many different ways to format data summary tables. Adding p-values to a table is just one thing that can be done in more than one way. For example, if a row group reports the counts and percentages for each level of a categorical variable across multiple (column) groups, then I would argue that the p-value resulting from a chi square test or a Fisher exact test would be best placed on the line of the table labeling the row group. However, say we reported the minimum, median, mean, and maximum with in a row group for one variable. The p-value from a t-test, or other meaningful test, for the difference in mean I would suggest should be reported on the line of the summary table for the mean, not the row group itself.

With so many possibilities I have reserved construction of a p-value column to be ad hoc. Perhaps an additional column wouldn’t be used and the p-values are edited into row group labels, for example.

If you want to add a p-value column to a qwraps2_summary_table object you can with some degree of ease. Note that qwraps2_summary_table objects are just character matrices.

See Table \@ref(tab:pValue).


```{r}
both %>% str
```


```{r}
pvals <-
  list(lm(mpg ~ am:vs,  data = mtcars2),
       lm(disp ~ am:vs, data = mtcars2),
       lm(disp ~ am:vs, data = mtcars2),  # yeah, silly example this is needed twice
       lm(wt ~ am:vs,   data = mtcars2)) %>%
  lapply(aov) %>%
  lapply(summary) %>%
  lapply(function(x) x[[1]][["Pr(>F)"]][1]) %>%
  lapply(frmtp) %>%
  do.call(c, .)
pvals
```

Adding the p-value column is done as follows:

```{r}
both <- cbind(both, "P-value" = "")
both[grepl("mean \\(sd\\)", rownames(both)), "P-value"] <- pvals
```

and the resulting table is:

```{r}
both
```

Table: (\#tab:pValue) There are many, many different ways to format data summary tables. Adding p-values to a table is just one thing that can be done in more than one way.



