---
title: "R Notebook"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: yes
---

https://mran.revolutionanalytics.com/snapshot/2018-01-16/web/packages/summarytools/vignettes/Introduction.html


```{r setup, include=FALSE, error=TRUE, message=FALSE, warning=FALSE} 
knitr::opts_chunk$set(echo=TRUE, 
                      comment=NA, 
                      error=TRUE, 
                      warning=FALSE, 
                      message = FALSE, 
                      results = "asis",
                      fig.align = 'center')
```


```{r}
library(summarytools)
```



```{r}
data(tobacco) # tobacco is an example dataframe included in the package
```



## with, by, view

Refer to Table \@ref(tab:usingBy)

```{r usingBy}
BMI_by_age <- with(tobacco, 
                   by(BMI, age.gr, descr, 
                      stats = c("mean", "sd", "min", "med", "max"), 
                      caption = "(\\#tab:usingBy) Using by()"
                      )
                   )

view(BMI_by_age, "pander", style = "rmarkdown",
     caption = "(\\#tab:usingBy) Using by()",
     omit.headings = TRUE)
```

Table: (\#tab:usingBy) Variable descriptive statistics. BMI by age.




## freq
See Table \@ref(tab:byGender)

```{r byGender}
freq(tobacco$gender, style="rmarkdown", 
     caption = "(\\#tab:byGender) Tobacco by Gender", 
     omit.headings = TRUE)
```


## Descriptive Statistics

### descr() with caption inside
See Table \@ref(tab:descrInside). It will not work.

```{r descrInside}
descr(tobacco, style = 'rmarkdown', 
      # caption = "(\\#tab:descrInside) Tobacco by Gender", 
      omit.headings = TRUE)
```

Table: (\#tab:descrInside) Tobacco by Gender



### descr() with caption outside
Refer to Table \@ref(tab:descrOutside).

```{r descrOutside}
descr(tobacco, style = 'rmarkdown', 
      omit.headings = TRUE)
```

Table: (\#tab:descrOutside) Dataframe descriptive statistics


## with(), by()
Table \@ref(tab:withTobacco)

```{r withTobacco}
with(tobacco, view(by(diseased, smoker, freq), method = "pander", 
                   # caption = "(\\#tab:withTobacco) Tobacco frequency by diseased, smoker",
                   omit.headings = TRUE))
```

Table: (\#tab:withTobacco) Tobacco frequency by diseased, smoker


## `Table` reference below chunk     
Table \@ref(tab:withTobaccoLabelDown)

```{r withTobaccoLabelDown}
with(tobacco, view(by(diseased, smoker, freq), method = "pander", 
                   omit.headings = TRUE))
```

Table: (\#tab:withTobaccoLabelDown) Tobacco. No smokers.



   



