---
title: "R Notebook"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
    toc: yes
---

https://mran.revolutionanalytics.com/snapshot/2018-01-16/web/packages/summarytools/vignettes/Introduction.html


```{r setup, include=FALSE, error=TRUE, message=FALSE, warning=FALSE} 
knitr::opts_chunk$set(echo=TRUE, 
                      comment=NA, 
                      error=TRUE, 
                      warning=FALSE, 
                      message = FALSE, 
                      results = "asis",
                      fig.align = 'center')
```


```{r}
library(summarytools)
```


```{r}
#st_options('omit.headings', TRUE)
st_options('bootstrap.css', FALSE)
st_options('footnote', NA)
```

```{r}
names(tobacco)
```

## freq

```{r}
freq(tobacco$gender, style = 'rmarkdown', 
     caption = "Tobacco by Gender")
```

## Cross-Tabulation


```{r}
ctable(tobacco$gender, tobacco$smoker, style = 'rmarkdown', 
       caption = "Tobacco. Cross-Tabulation by Gender (rolling percentage)", 
       omit.headings = TRUE)
```

### Descriptive Statistics
See Table \@ref(tab:descrTobacco)

```{r}
descr(tobacco, style = 'rmarkdown', 
      caption = "Tobacco. Descriptive Statistics")
```

Table: (\#tab:descrTobacco) Tobacco. Descriptive Statistics



# Data frame Summaries

## grid, no plain ascii
See Table \@ref(tab:dfSummaryGrid)

```{r}
dfSummary(tobacco, style = 'grid', plain.ascii = FALSE,
          caption = "Tobacco. Dataframe Summary")
```

Table: (\#tab:dfSummaryGrid) Tobacco. Dataframe Summary with graph


## no graph column, no plain ascii
See Table \@ref(tab:dfSummaryNoGraph)

```{r}
dfSummary(tobacco, style='grid', plain.ascii = FALSE, graph.col = FALSE)
```

Table: (\#tab:dfSummaryNoGraph) Dataframe Summary with no graph


## Using by() to Show Results By Groups
See Table \@ref(tab:usingBy)

```{r}
# First save the results
iris_stats_by_species <- by(data = iris, 
                            INDICES = iris$Species, 
                            FUN = descr, stats = c("mean", "sd", "min", "med", "max"), 
                            transpose = TRUE,
                            caption = "Using by()")

# Then use view(), like so:
# view(iris_stats_by_species, method = "pander", style = "rmarkdown")
view(iris_stats_by_species, method = "pander", style = "rmarkdown",
    caption = "Using by()")
```

Table: (\#tab:usingBy) iris stats by species.


## Special Case - Using descr() With by() For A Single Variable
Refer to Table \@ref(tab:with-by)

```{r usingBy}
data(tobacco) # tobacco is an example dataframe included in the package
BMI_by_age <- with(tobacco, 
                   by(BMI, age.gr, descr, 
                      stats = c("mean", "sd", "min", "med", "max"), 
                      caption = "Using by()"))

view(BMI_by_age, "pander", style = "rmarkdown",
     caption = "Using by()",
     omit.headings = TRUE)
```

Table: (\#tab:with-by) BMI by ages. Using `with()` and `by()`. Single Variable.


### BMI_by_age is not a dataframe

```{r}
class(BMI_by_age)
```

## Transpose

The transposed version looks like this:
Refer to Table \@ref(tab:transposed)

```{r}
BMI_by_age <- with(tobacco, 
                   by(BMI, age.gr, descr,  transpose = TRUE,
                      stats = c("mean", "sd", "min", "med", "max")),
                   caption = "Transposed Descriptive")

view(BMI_by_age, "pander", style = "rmarkdown", omit.headings = TRUE,
     caption = "Transposed Descriptive")
```

Table: (\#tab:transposed) BMI by ages. Single Variable, transposed.


## Using lapply() to Show Several freq() tables at once
Refer to Table \@ref(tab:lapplyFreqTables)

```{r}
tobacco_subset <- tobacco[ ,c("gender", "age.gr", "smoker")]
freq_tables <- lapply(tobacco_subset, freq)
# view(freq_tables, footnote = NA, file = 'freq-tables.html')
view(freq_tables, method = 'pander', omit.headings = TRUE)
```

Table: (\#tab:lapplyFreqTables) Frequency Tables


# Using by(), with(), and lapply()
Refer to Table \@ref(tab:examsDescriptive)

```{r}
stats <- by(data = exams$geography, INDICES = exams$gender, FUN = descr, 
            style = 'rmarkdown')

view(stats, method = 'pander', omit.headings = TRUE)
```

Table: (\#tab:examsDescriptive) Descriptive Statistics for Geography class by Gender


```{r results="markup"}
what.is(iris)
```


